Drivers of tree regeneration remain challenging to be captured by models of forest dynamics

### 2022-11-21 manuscript/manuscriptRW.Rmd
updates in dicusssion (still very messy)

---
title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries = FALSE)
  library(bibtex)
  #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")

@@

Tree recruitment depends on multiple processes such as reproduction,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival; these processes are all difficult to assess and
some of them are scarcely understood and thus appear highly stochastic.
Finding the right level of complexity in the definition of the factors
leading to tree recruitment is challenging and models of forest dynamics
have used different approaches [@bugmann2022]. Overall, models are
needed to (1) synthesize existing empirical data and explore their
consequences, and (2) assess future tree recruitment e.g. in the context
of global change scenarios. Currently, regeneration processes in dynamic
forest models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022], from entirely ignoring tree regeneration (as done in
classical forest growth models, e.g. @pretzsch2002), across the use of a
few simple environmental filters, as done in most forest \"gap\" models
(e.g., @shuman2011) and Dynamic Global Vegetation Models (e.g.,
@smith2001; @hickler2012), to complex approaches that incorporate local
feedback from the canopy, multiple ecological processes and often also
small time steps (e.g., @wehrli2006; @seidl2012), or field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time (e.g., @ribbens1994).

With the current strategies that models of forest dynamics are using to
represent tree recruitment we are obtaining results that faced several
problems related to over-recruitment, needed excess of mortality at
early stages and inadecuate species composition in regeneration which
have strong impacts on forest dynamics. Some model outputs are
over-recruiting which subsequently must die primarily by competition in
later stages. This is characteristic of \'classical\' forest gap models
[@gutiérrez2016], at least as long as the simulations comprise a
multitude of species. Other models use a model output to correct
recruitment levels for example by routinely calibrate them against the
predicted canopy species composition affecting as well the species
dominance.

Correctly capturing the species composition in recruitment is often also
a challenge, thus leading to unrealistic successional drift in the model
REFERENCE?, for example by the change of the dominant species via
stochastic processes that continuously bring seeds of other species that
the come to dominance, with this strategy a new species might start to
rise eliminating the former dominant species. A related issues is the
rarification of species richness due to positive feedback effects such
that eventually just single-species forest stands remain. This is
sometimes handled by the incorporation of a low level of seed influx of
all species at all times (e.g., @lischke2006 ). However, simulated
species composition is usually exceedingly sensitive to that level
whereas the parameters of such functions are poorly constrained by field
data. There is also often a problem with the insuficient constraint of
parameter values for models that start from very small tree sizes (e.g.,
10 cm of height or even seeds) and track tree development in a
process-oriented manner by considering a multitude of ecological
influences, rather than being statistically driven and emphasizing
recruitment into a larger size class (e.g., @wehrli2006 ). Most of these
undesired behaviors in tree recruitment are often not published as such,
but in the vast majority of studies they are corrected (e.g.,
@rasche2011 , where for most study sites regeneration was simply
switched off). Thus, a strong focus on the modeling of tree regeneration
processes is highly important and long overdue [@price] if we are to
make reliable projections of future forest dynamics, i.e. when the
models need to be operated in extrapolation mode, but also from a
fundamental ecological point of view (increasing systems understanding).

This is the first time, a large number of forest models commonly used to
assess forest dynamics under climate change are evaluated against a
continental-scale multi-species harmonized data set on tree recruitment
[@käber2022]. Tree recruitment is understudied in field studies and
modelling [@hanburybrown_forest_2022], because little data has been
available and data was very fragmented, i.e., not in one harmonized data
set. In this study we relied on models with various 'philosophies',
including models operating at the stand to the global scale. Because of
the large variability in recruitment patterns in nature and the large
amount of factor driving recruitment, including some which are not
included in most models, such as deer browsing, we do not aim at a
detailed statistical evaluation of each model. Instead, we aim at
evaluating general recruitment patterns, magnitudes and benchmark the
regeneration niche of multiple species based on empirical data along a
wide environmental gradient (temperature, moisture, and light
availability). More specifically in this study we wanted to answer the
following questions: 1) are models of forests dynamics capturing
correctly regeneration levels across environmental gradients; 2) are
models of forest dynamics representing correctly initial composition of
tree populations;and, 3) can models of forest dynamics correctly capture
the initial mortality in tree in-growth?.

@@

@@

**General**

Tree recruitment is a fundamental process in forest dynamics, in
particular with increasing disturbance under climate change.

-   The problem of correctly capturing post-disturbance recovery, how
    seed availability and climate conditions define the 15-20 years
    status of a forest. If this is not well capture by the models we can
    not really assess the impacts of disturbances on forests.

```{=html}
<!-- -->
```
-   the approach: the idea of confronting the simulated recruitment
    along environmental gradients with data from such gradients.

-   the idea of not doing a calibration a recruitment model based on the
    EuFoRIa data, and I explain why it is more insightful to use these
    data for model benchmarking: we can learn more

**Regeneration levels**

-   Implication of models that produce very high initial densities

-   Range shifts retractions in species expansions because of climate
    change might not also be properly represented if we do not capture
    properly recruitment rates in the environmental gradient .

-   importance of light availability for regeneration and the proxy of
    stand basal area to define that . Can we see a expected pattern that
    shows how recruitment of the different species is becoming sparser
    and drops out completely as a function of a proxy of light
    availability.

**Regeneration species composition**

-   If the composition is not well captured then we can not assess the
    functional composition of the forest appropriately, e.g. their
    sensitive to drought or resilience to new disturbances events.

-   Species richness variations might be explained by the way each model
    considers the recruitment (e.g. in the absence of adults recruitment
    can be linked to either seed rain form elsewhere -dispersal, or to
    sees/seedling bank) Differences might also be expected between
    cohort and individual trees based models.

@@

-   Is the adult forest the right forest, how is this relevant and is
    this the consequence for the correct recruitment composition? This
    is highly dependent on the model structure:

    -   Background seed Yes/no

    -   Stochastic vs deterministic

@@

-   Most models maintain or even increase in recruitment diversity
    between the 7 and 10 cm thresholds, and the same is visible from the
    observed data, however the differences are not significant. EXPLAIN
    WHY This cannot be species richness (as there are no new species
    popping between 7 and 10 cm DBH). I'd also expect the number of
    species to be lower at 10 cm than at 7 cm (because some will die off
    due to insufficient light etc.), but this is not reflected in H'
    because species with zero abundance are ignored (ln[0] is not
    defined). So a strongly increasing evenness must be driving this
    increase of H'." This could be because we are comparing 7 and 10cm
    from same samples, and at 10cm you might include more trees and at
    7cm This calls for further research and another paper?

**Regeneration mortality**

-   Implications of models that have lower mortality in regeneration but
    large recruitment

-   We expected that the higher regeneration at 7 cm the stronger the
    regeneration mortality (decrease in stem number between regeneration
    at 7cm and regeneration at 10cm) . Not the case, why, implications

-   Models below the Reineke line have lower mortality than expected by
    the self-thinning rule, which may point to facilitation, or simply a
    higher mortality rate before the trees reached 7 cm diameter. But
    then these models would need to underestimate recruitment levels
    (Fig. 1), no? But this is not what we see... not sure how to
    explain.

-   There might be a difference between models including stochastic
    processes in recruitment and those which do not, but perhaps the
    reality implies higher mortality rates in the adult component
    (enhance the share of less shade tolerant species?)

-   Comments on the Reineke rule

@@

**Zero inflated**

-   Observed data has much more sites without any recruitment (zero
    inflated data). Many models seem to predict recruitment in all
    sites. The zeros in the empirical data come from no recruitment and
    from thresholds larger than 10cm for some sites. Is this a
    consequence of assuming that the observed data is also in
    equilibrium. Lack of recruitment in sites an indication of
    "non-equilibrium" due to management history?

-   the reason why some models have zero regeneration at very low BA
    values may be that they never actually simulate such low BA values
    at the one ha level (which was the sampling unit if I recall
    correctly). If so, this should be pointed out as well, I think.

**Other aspects**

@@

**Models structure and recomendations**

>>>>

---
title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  word_document:
    reference_docx: template_paper.docx
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries = FALSE)
  library(bibtex)
  options(tinytex.verbose = TRUE)
  #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")

@@

Tree recruitment depends on multiple processes such as reproduction,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival; these processes are all difficult to assess and
some of them are scarcely understood and thus appear highly stochastic
[@bogdziewicz2021]. Finding the right level of complexity in the
definition of the factors leading to tree recruitment is challenging and
models of forest dynamics have used different approaches [@bugmann2022].
Overall, models are needed to (1) synthesize existing empirical data and
explore their consequences, and (2) assess future tree recruitment e.g.
in the context of global change scenarios. Currently, regeneration
processes in dynamic forest models are handled in a multitude of ways
[@konig_tree_2022; @bugmann2022], from entirely ignoring tree
regeneration (as done in classical forest growth models, e.g.
@pretzsch2002), across the use of a few simple environmental filters, as
done in most forest \"gap\" models (e.g., @shuman2011) and Dynamic
Global Vegetation Models (e.g., @smith2001; @hickler2012), to complex
approaches that incorporate local feedback from the canopy, multiple
ecological processes and often also small time steps (e.g., @wehrli2006;
@seidl2012), or field-based statistical parameterizations, which however
are not easy to extrapolate in space and time (e.g., @ribbens1994).

With the current strategies that models of forest dynamics are using to
represent tree recruitment we are obtaining results that faced several
problems and have strong impacts on forest dynamic such as
over-recruitment, needed excess of mortality at early stages and
inadecuate species composition in regeneration. Some model outputs are
over-recruiting which subsequently must die primarily by competition in
later stages. This is characteristic of \'classical\' forest gap models
[@gutiérrez2016], at least as long as the simulations comprise a
multitude of species. Other models use a model output to correct
recruitment levels for example by routinely calibrate them against the
predicted canopy species composition affecting as well the species
dominance.

Correctly capturing the species composition in recruitment is often also
a challenge, thus leading to unrealistic successional drift in the model
REFERENCE?, for example by the change of the dominant species via
stochastic processes that continuously bring seeds of other species that
the come to dominance, with this strategy a new species might start to
rise eliminating the former dominant species. A related issues is the
rarification of species richness due to positive feedback effects such
that eventually just single-species forest stands remain. This is
sometimes handled by the incorporation of a low level of seed influx of
all species at all times (e.g., @lischke2006 ). However, simulated
species composition is usually exceedingly sensitive to that level
whereas the parameters of such functions are poorly constrained by field
data. There is also often a problem with the insuficient constraint of
parameter values for models that start from very small tree sizes (e.g.,
10 cm of height or even seeds) and track tree development in a
process-oriented manner by considering a multitude of ecological
influences, rather than being statistically driven and emphasizing
recruitment into a larger size class (e.g., @wehrli2006 ). Most of these
undesired behaviors in tree recruitment are often not published, but in
the vast majority of studies they are corrected (e.g., @rasche2011,
where for most study sites regeneration was simply switched off). Thus,
a strong focus on the modeling of tree regeneration processes is highly
important and long overdue [@price] if we are to make reliable
projections of future forest dynamics, i.e. when the models need to be
operated in extrapolation mode, but also from a fundamental ecological
point of view (increasing systems understanding).

Tree recruitment is understudied in field studies and modelling
[@hanburybrown_forest_2022], as little data has been available and it is
often very fragmented, i.e., not in one harmonized data set. On the one
hand, monitoring on permanent sample plots (such as national forest
inventories) often measure tree recruitment [@zell_modeling_2019]
however often with a design that captures the regeneration intensity
after the stand initiation phase on productive forests [@hallsby2015].
Data from permanent plots, although available across regions, it is also
highly heterogeneous in design making their usability more difficult
[@käber2021]. On the other hand, targeted studies to measure tree
recruitment on experimental sites are often limited in space, species or
the relevant processes leading to regeneration that are targeted
[@berdanier2016; @collet2006].

This is the first time, a large number of forest models commonly used to
assess forest dynamics under climate change are evaluated against a
continental-scale multi-species harmonized data set on tree recruitment
[@käber2022]. In this study we relied on models with various
'philosophies', including models operating at the stand to the global
scale. Because of the large variability in recruitment patterns in
nature and the large amount of factor driving recruitment, including
some which are not included in most models, such as deer browsing, we do
not aim at a detailed statistical evaluation of each model. Instead, we
aim at evaluating general recruitment patterns, magnitudes and benchmark
the regeneration niche of multiple species based on empirical data along
a wide environmental gradient (temperature, moisture, and light
availability). More specifically in this study we wanted to answer the
following questions: 1) are models of forests dynamics capturing
correctly regeneration levels across environmental gradients; 2) are
models of forest dynamics representing correctly initial composition of
tree populations;and, 3) can models of forest dynamics correctly capture
the initial mortality in tree in-growth?.

@@

Two models, ForClim and ForCEEPs used two alternative model versions.
HERE **FORCLIM** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCLIM 1
AND 11 AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS. HERE
**FORCEEPS** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCEEPS AND
FORCEEPS(F) AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS.

@@

Tree recruitment is a fundamental process in forest dynamics, in
particular with increasing disturbance under climate change . Correctly
capturing regeneration in forest dynamics is fundamental to evaluate for
example post-disturbance dynamics as this will define the status of the
forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
initial state it is not well captured then we can not properly assess
aspects such as the forest resilience, or changes in carbon capture. In
this study, for the first time, projections of tree recruitment done by
multitude models of forest dynamics were confronted with observed data
across a large environmental gradients. The observed data that we used
[@käber2022] is exceptional in this regard, and that's why it was used
for model benchmarking rather than calibration.

Adequate tree recruitment rates define the structure and composition of
forests. Most of the models of forest dynamics tended to over-recruit
trees in all the sites. Excess in recruitment has to lead to a large
mortality in later stages. Capture properly recruitment rates in the
environmental gradient might impact future forests. If not capture well
it is going to be difficult to for models to correctly project range
shifts retractions in species expansions because of climate change might
not also be properly represented if we do not capture properly
recruitment rates in the environmental gradient.

Regeneration levels define the success of future forests, but it is
equally important to correctly identify un-stocked forest areas
[@rammer2021]. Tree recruitment is often characterized by an excess
number of zero counts [@fortin2007]. In this study some of the models
did not have lack of recruitment or their proportion was very low. This
substantial difference with the observed data might be a consequence of
assuming that the observed data is also in equilibrium, where lack of
recruitment in sites could also be an indication of "non-equilibrium"
due to management history.

Light availability is an important process defining tree recruitment
[@collet2006; @berdanier2016]. However it is difficult to have measured
variables defining light availability for the regeneration. In this
study we used the total stand basal area as a proxy for light
availability. There was differences in the basal area ranges, some of
the models had stand basal area lower than the observe data but there
was also models that did have more stocked stands that the observed data
and therefore it was not possible to evaluate the recruitment for lower
density stands. It was also not possible to observed the expected
pattern showing how recruitment of the different species is becoming
sparser and drops out completely as a function of a proxy of light
availability.

If the recruitment tree composition is not well captured then we can not
assess the functional composition of the forest appropriately, e.g.
their sensitive to drought or resilience to new disturbances events. In
the models simulation we saw variations on how the main tree species
were represented in the environmental gradient but also in the in
overall species richness levels. Most models maintain or even increase
in recruitment diversity between the 7 and 10 cm thresholds, and the
same is visible from the observed data, however the differences are not
significant. Species richness variations might be explained by the way
each model considers recruitment, e.g. in the absence of adults
recruitment can be linked to either seed rain from elsewhere -dispersal,
or to seed/seedling bank). Differences might also be expected between
cohort and individual trees based models.

@@

-   Is the adult forest the right forest, how is this relevant and is
    this the consequence for the correct recruitment composition? This
    is highly dependent on the model structure:Background seed Yes/no /
    Stochastic vs deterministic

@@

We could expect that models that over-recruited trees. will also have a
large mortality between 7 and 10 cm thresholds. This implies that this
models have to have the excess of mortality in later stages of the
forest dynamics. Models below the Reineke line have lower mortality than
expected by the self-thinning rule, which may point to facilitation, or
simply a higher mortality rate before the trees reached 7 cm diameter,
however one would also assume that these models will underestimate
recruitment which was not the case. Why?. There might be a difference
between models including stochastic processes in recruitment and those
which do not, but perhaps the reality implies higher mortality rates in
the adult component (enhance the share of less shade tolerant species).

There is also the question of how much level of detail one have to give
to define regeneration processes [@konig_tree_2022; @bugmann2022].
Models with more detailed and complex recruitment modules, will have
also increase parametrisation which will lead to lower generalization as
this values will have to be collected from specific geographical
regions.

@@

-   We use all the models to rpoduce values per 1ha, stand ang global
    models potentially disregarding the advantages that landscape models
    present with spatial seed dispersal (TreMig, iLand and LAndClim
    references). Without spatial context we are limited to the seed
    influx from the stand itself if the model does not have a background
    seed input. But even global vegetation models are lacking this
    dispersal between cells [@hanburybrown_forest_2022] abstraction of
    space or even the trees themselves. It is important to include
    spatial considerations on tree recruitmen [@beckage2003]

-   Comparing PFTs with species. We lose the species specific
    information on with substantial drawbacks [@purves2008]

-   The loss of species-specific environmental responses through the
    collation of species communities into PFTs has been widely accepted,
    as parameterization efforts would otherwise exceed the available
    means. However, it is doubtful whether the definition of PFTs
    provides sufficient flexibility for an adequate representation of
    the ecological processes and the differences between species

-   Recruitment of trees into size classes of ≥4 cm DBH, as in our four
    data sets, represents a relatively late stage of forest
    regeneration. It integrates over sometimes long periods since
    germination and multiple environmental constraints act upon young
    trees of widely different size. kaber 2021

@@

# Data availability statement

The code and data that support the findings of this study are openly
available at zenodo and gitlab (links to be added)
